---
title: "Dose-response meta-analysis: Brexpiprazole example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dose-response meta-analysis: Brexpiprazole example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 6,
  fig.height = 4
)
```

## Overview

This vignette reproduces the primary dose-response analysis from:

> Furukawa Y, Oguro S, Obata S, Hamza T, Ostinelli EG, Kasai K.
> Optimal dose of brexpiprazole for augmentation therapy of
> antidepressant-refractory depression: A systematic review and
> dose-effect meta-analysis.
> *Psychiatry Clin Neurosci*. 2022;76(9):416–422.
> doi:[10.1111/pcn.13438](https://doi.org/10.1111/pcn.13438)

The underlying statistical method is the one-stage dose-response meta-analysis
described in:

> Crippa A, Discacciati A, Bottai M, Spiegelman D, Orsini N.
> One-stage dose-response meta-analysis for aggregated data.
> *Stat Methods Med Res*. 2019;28(5):1579–1596.
> doi:[10.1177/0962280218773122](https://doi.org/10.1177/0962280218773122)

`drma` is a wrapper around the
[`dosresmeta`](https://cran.r-project.org/package=dosresmeta) package that
implements this approach.  **Please cite both papers when using `drma`.**

---

## Data

The `brexpiprazole` dataset (14 arms, 6 trials) is bundled with the package.

```{r data}
library(drma)
data(brexpiprazole)
brexpiprazole[, c("study_id", "dose", "n_arm", "n_responders")]
```

---

## 1. Efficacy (response rate, OR)

Pass arm-level event counts; `drma` computes the log-OR internally.

```{r efficacy}
res_e <- drma(
  data    = brexpiprazole,
  studlab = study_id,
  dose    = dose,
  sm      = "OR",
  event   = n_responders,
  n       = n_arm,
  knots   = c(1, 2, 3)
)
print(res_e)
```

```{r plot-efficacy}
plot(res_e,
     ylab     = "Response (OR)",
     ylim     = c(0.75, 2),
     xlab     = "Brexpiprazole (mg)",
     ref_dose = 0,
     bubble   = TRUE,
     rug      = TRUE)
```

---

## 2. Tolerability and acceptability (precomputed log-OR)

When log-ORs are already computed (e.g. from a published table), use
`sm = "precomputed"`.

```{r tolerability}
res_t <- drma(
  data    = brexpiprazole,
  studlab = study_id,
  dose    = dose,
  sm      = "precomputed",
  yi      = tolerability_logor,
  sei     = tolerability_se,
  event   = n_dropout_ae,
  n       = n_arm,
  knots   = c(1, 2, 3)
)

res_a <- drma(
  data    = brexpiprazole,
  studlab = study_id,
  dose    = dose,
  sm      = "precomputed",
  yi      = acceptability_logor,
  sei     = acceptability_se,
  event   = n_dropout_any,
  n       = n_arm,
  knots   = c(1, 2, 3)
)
```

```{r plot-tol-acc, fig.width = 8}
par(mfrow = c(1, 2))
plot(res_t, ylab = "Dropout for AEs (OR)",         ylim = c(0.2, 5),
     xlab = "Brexpiprazole (mg)", ref_dose = 0, bubble = TRUE, rug = TRUE)
plot(res_a, ylab = "Dropout for any reason (OR)",  ylim = c(0.2, 5),
     xlab = "Brexpiprazole (mg)", ref_dose = 0, bubble = TRUE, rug = TRUE)
par(mfrow = c(1, 1))
```

---

## 3. Sensitivity analyses — knot placement

```{r sensitivity}
res_s1 <- drma(
  data = brexpiprazole, studlab = study_id, dose = dose, sm = "OR",
  event = n_responders, n = n_arm, knots = c(0.5, 1.5, 2.5)
)
res_s2 <- drma(
  data = brexpiprazole, studlab = study_id, dose = dose, sm = "OR",
  event = n_responders, n = n_arm, knots = "0.25-0.50-0.75"
)
res_s3 <- drma(
  data = brexpiprazole, studlab = study_id, dose = dose, sm = "OR",
  event = n_responders, n = n_arm, knots = "0.1-0.5-0.9"
)
res_s4 <- drma(
  data = brexpiprazole, studlab = study_id, dose = dose, sm = "OR",
  event = n_responders, n = n_arm, curve = "log"
)
```

```{r plot-sensitivity}
plot(res_e,  col = "black",       lwd = 2,
     ylim = c(0.75, 2), ylab = "Response (OR)",
     xlab = "Brexpiprazole (mg)", ref_dose = 0)
lines(res_s1, col = "tomato",      lty = 2, lwd = 1.5)
lines(res_s2, col = "steelblue",   lty = 3, lwd = 1.5)
lines(res_s3, col = "forestgreen", lty = 4, lwd = 1.5)
lines(res_s4, col = "purple",      lty = 5, lwd = 1.5)
legend("topleft",
       legend = c("Primary c(1,2,3)", "S1 c(0.5,1.5,2.5)",
                  "S2 25/50/75th pct", "S3 10/50/90th pct", "S4 log-linear"),
       col = c("black", "tomato", "steelblue", "forestgreen", "purple"),
       lty = 1:5, lwd = 1.5, bty = "n")
```

---

## 4. Predictions at clinically relevant doses

`baseline_prop = 0.183` converts OR to absolute response rate assuming an
18.3% baseline response (placebo arm pooled proportion from the paper).

```{r predict}
predict_table(
  res_e,
  doses         = c(0, 0.5, 1, 1.5, 2, 3),
  ref_dose      = 0,
  baseline_prop = 0.183
)
```

---

## 5. Target doses (ED50 / ED95 / ED100)

```{r target}
target_dose(res_e, p = c(0.5, 0.95, 1))
```

---

## 6. League table

Pairwise OR between every combination of doses.

```{r league}
league_table(res_e, doses = c(0, 1, 2, 3))
```

---

## Citation

```r
citation("drma")
```

When reporting results, cite the one-stage dose-response method:

> Crippa A, Discacciati A, Bottai M, Spiegelman D, Orsini N.
> One-stage dose-response meta-analysis for aggregated data.
> *Stat Methods Med Res*. 2019;28(5):1579–1596.
> doi:10.1177/0962280218773122

and the `dosresmeta` package:

> Crippa A, Orsini N.
> Dose-response meta-analysis of differences in means.
> *BMC Med Res Methodol*. 2016;16:91.
> doi:10.1186/s12874-016-0189-0
